Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-13 WSSSPE1_report/WSSSPE1_summary_report.tex
Section 3:  Keynote: Add line breaks.
Section 3.2:  Keynote 2 strengthened and lengthened.

C. Titus Brown~\cite{ged-web}, a WSSSPE1 contributor and participant, argued that with regard to reusable
software, ``we should just start doing it.'' Smith put forward that
documentation should be ``treated as a first class entity.''  He noted
that the open source community has excellent cultures of code reuse, for example,
RubyGems~\cite{rubygems-web}, PyPI~\cite{pypi-web}, and
CPAN~\cite{cpan-web}, where there is effectively low-friction
collaboration through the use of repositories. This has not happened in
highly numerical, compiled language scientific software.
An exception he cited as a good example of scientific
projects using GitHub is the EMCEE Markov Chain Monte Carlo project~\cite{emcee-web} by
Dan Foreman-Mackey and contributors.

He argued that the GitHub Pull Request code review mechanism
facilitates such collaboration, by allowing one to code first, and
seek review and a merge back into the trunk later.

Fernando Perez from the audience emphasized that ``open source is reproducible by
necessity'' - since reproducibility is a prerequisite for remote
collaboration. In response, Smith defined GitHub as part of the next
stage in web development, following on from the social web of
Facebook, i.e., ``The Collaborative Web.''

>>>>

C. Titus Brown~\cite{ged-web}, a WSSSPE1 contributor and participant,
argued that with regard to reusable software, ``we should just start
doing it.'' Smith replied that documentation should be ``treated
as a first class entity.''  He noted that the open source community
has excellent cultures of code reuse, for example,
RubyGems~\cite{rubygems-web}, PyPI~\cite{pypi-web}, and
CPAN~\cite{cpan-web}, where there is effectively low-friction
collaboration through the use of repositories. This has not happened
in highly numerical, compiled language scientific software.  An
exception he cited as a good example of scientific projects using
GitHub is the EMCEE Markov Chain Monte Carlo project~\cite{emcee-web}
by Dan Foreman-Mackey and contributors.

He argued that  GitHub's \emph{Pull Request} code review mechanism
facilitates such collaboration, by allowing one to code first, and
seek review and merge back into the trunk later.

``Open source is \ldots reproducible by necessity,''
Smith quoted  Fernando Perez~\cite{perez-open-src-reproducible}, explaining that 
reproducibility is a prerequisite for remote collaboration.  He pointed out that GitHub could be the propeller
of the next stage in web development, i.e., ``the Collaborative Web,'' following on from the social web of
Facebook.

In conclusion Smith reiterated the importance of establishing incentive models 
for open contributions and tool builders, for example, meaningful metrics 
and research grants such as \cite{NSF_software_vision}. He urged computational scientists to 
collabroate and share often  their research reports, teaching materials, 
code, as well as data by attaching proper licenses.
